---
title: "greenbuildings_analysis"
author: "Meenakshi Sundaram (UT EID ms88544), Anisha Singh(UT EID as229583),Antra Sinha(UT EID as224589),Hayong Kim(UT EID hk26786)"
date: "2023-08-08"
output:
  html_document: default
---

**Q3 : Visual story telling part 1: Green Buildings:**

```{r,message=FALSE,warning=FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)
library(viridis)
library(kableExtra)
library(plotly)

data <- read.csv("C:\\Users\\kowsh\\OneDrive\\Documents\\greenbuildings.csv",header=TRUE)

data$green_rating=factor(data$green_rating)
data$class_a=factor(data$class_a)

summary_table <- data %>% 
  group_by(green_rating,class_a) %>% 
  summarize(n_pct =round((n()/nrow(data))*100,3), 
            Med_Rent = median(Rent),
            Med_occupancy_rate=median(leasing_rate),
            Med_age=median(age)) %>% 
  arrange(desc(Med_Rent)) 

#formating the table using kableExtra library
format_summary_table <- summary_table%>% 
  kable(format="html",escape=FALSE) %>% 
  kable_styling(full_width=FALSE) %>% 
  row_spec(1,background = "#FFFF99")

format_summary_table

#formatting the table for plotting a horizontal bar chart
summary_data_long <- summary_table %>%
  pivot_longer(cols = c(Med_Rent, Med_age, Med_occupancy_rate),
               names_to = "variable",
               values_to = "value")

p <- ggplot(summary_data_long, aes(x = interaction(green_rating, class_a), y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = value), position = position_dodge(width = 1),hjust=1.1, vjust = 0.5, color = "black") +
  labs(
    x = "Green Rating & Class A",
    y = "Values",
    title = "Median Rent, Median Age, and Median Occupancy by Green Rating and Class A"
  ) +
scale_fill_manual(values = c(
    "Med_Rent" = "#66c2a5",
    "Med_age" = "#fc8d62",
    "Med_occupancy_rate" = "#8da0cb")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  coord_flip()

p

green_color <- "#1B9E77"  #color for green buildings
non_green_color <- "#D95F02"  #color for non-green buildings


#creating faceted histogram for green_rating by class
ggplot(data)+
  geom_bar(aes(x=green_rating,fill=class_a))+
  scale_fill_brewer(type="qual")+
  facet_wrap(~class_a)+
  geom_text(aes(x = green_rating, label = scales::percent(..count../sum(..count..))),hjust=0.4,
  stat = "count",
  position = position_stack(vjust = 0.9),
  show.legend = FALSE,
  size = 3)

#visualizing the data to identify confounding variables affecting rent

#scatter plot of size vs rent
ggplotly(ggplot(data)+
  geom_point(aes(x=size, y=Rent,color=green_rating),alpha=0.7)+
  scale_color_manual(values=c(non_green_color,green_color))+
  labs(x="Size in sqft",
       y="Rent in $/sqft",
       title="Size Vs Rent",
       colour ="Is_green_building"))


#scatter plot of cluster rent vs rent
ggplotly(ggplot(data)+
  geom_point(aes(x=cluster_rent,y=Rent,color=green_rating),alpha=0.7)+
 scale_color_manual(values=c(non_green_color,green_color))+
  labs(x="Cluster Rent ($/sqft)",
       y="Rent ($/sqft)",
       title="Cluster Rent vs Rent",
       colour="Is_green_building"))

#box plot of class vs age by green_rating
ggplot(data)+
  geom_boxplot(aes(x=class_a,y=age,color=green_rating))+
  labs(x="Is_Class_A",
       y="Age of Building in Yrs",
       title="Box plot of Class vs Age by green_rating",
       colour="Is_green_building")+
  scale_color_viridis(discrete=T)


#age vs rent

ggplotly(ggplot(data)+
  geom_point(aes(x=age,y=Rent,color=green_rating),alpha=0.7)+
  scale_color_manual(values=c(non_green_color,green_color))+
  labs(x="Age in Years",
       y="Rent in $/Sqft",
       title=" Age vs Rent",
       colour="Is_green_building"))


#age vs rent faceted by class of buildings


ggplot(data,aes(x=age, y=Rent,color=class_a),alpha=0.6)+
  geom_point(aes(shape=class_a),alpha=0.7)+
  facet_wrap(~green_rating,labeller = labeller(green_rating = c("1" = "Green Buildings ", "0" = "Non-Green Buildings")))+
 scale_color_manual(values=c(non_green_color,green_color))
  

```
Observations : 

* Rent is positively correlated with size 

* Cluster rent and rent are positively correlated

* Green class A buildings are younger, have a higher rent and occupancy rate than other buildings

* Most of the class A buildings are younger, but irrespective of their age, they have a higher rent per sqft because of their premium building quality/market positioning

*The proportion of class A buildings is higher for green buildings. ~80% of the green buildings are class A buildings. 

* Also, the rent for class_a buildings are substantially higher than that of non-class_a buildings (~$4.8/sqft)

* There is no correlation between age and rent. Hence, we will exclude this variable from our analysis


```{r,message=FALSE,warning=FALSE}

library(plotly)

green_color <- "#1B9E77"  # Color for green buildings
non_green_color <- "#D95F02"  # Color for non-green buildings


size_bins <- seq(0, 2000000, by = 50000)

# Create bins using cut() function
data$size_groups <- cut(data$size, breaks = size_bins, labels = FALSE, right = FALSE)

clean_data=na.omit(data)

#density histogram of size

ggplotly(ggplot(clean_data) +
  geom_density(aes(x=size_groups,fill=green_rating),alpha=0.6)+
  facet_wrap(~class_a,labeller = labeller(class_a = c("1" = "Class A Buildings ", "0" = "Non- Class A Buildings")))+
    labs(x="size groups of width 50k sqft",
         y="Density of buildings for each size_group",
         title="Density histogram of sise groups",
         colour="Is_green_rating"))



Median <- clean_data %>% group_by(size_groups,green_rating,class_a) %>% summarize(n=n(),Med_rent=median(Rent,na.rm=TRUE))


plot <- ggplot(Median)+
  geom_line(aes(x=size_groups,y=Med_rent,color=green_rating),size=1.2)+
  facet_wrap(~class_a,labeller=labeller(class_a=c("1"="Class A Buildings","0"="Non-Class A Buildings")))+
    scale_color_manual(values=c(non_green_color,green_color))+
  scale_x_continuous(breaks=seq(2,40,by=2),labels = seq(2,40,by=2))+
  labs(x="Size Groups of Width 100k Sqft",
       y="Median Rent per Size Group in $/sqft",
       title = "Size Vs Rent by Green_rating and Class_a",
       colour="Is_green_building")+
  theme(axis.text.x=element_text(angle=45,hjust=1))

interactive_plot <- ggplotly(plot)

interactive_plot

Result_table <- 
  clean_data %>% 
  filter(size_groups==6) %>% 
  group_by(size_groups,green_rating,class_a) %>% 
  summarize(Med_rent=round(median(Rent),2),
            Med_occupancy = round(median(leasing_rate),1),
            Min_size=min(size)) %>% 
  arrange(desc(Med_rent))

  Format_Result_table <- Result_table %>%
    kable(format="html",escape=FALSE) %>% 
    kable_styling(full_width=FALSE) %>% 
    row_spec(c(1),background = "#FFFF99") %>% 
    row_spec(c(2),background = "#FF9999")
  
  Format_Result_table
  
  

```
Insights: 


* There seems to be an almost similar size distribution for both green and non-green buildings within the same class. Hence, we could use these splits to study the trend of median rent in these groups. 

*It is not advisable to invest in green buildings if its not a class a building as the rent is lower in this case when compared non-green buildings

*For a green, class A building of size between 250000 and 350000 sq ft, the median rents are higher than that of its non-green class A counterparts.  


Will investing in green buildings be worth it? 

Baseline construction costs + green rating premium = $105,000,000 


Baseline construction cost/sqft = $100 M/250K sqft =$400/sqft

Hence, going for bigger size buildings will increase the bottom line by a substantial margin and we therefore eliminate going for green, class A buildings of size 300k to 350k (group 7) from our recommendations. 


***Final Recommendation : Building a green, class A building of size 250k to 300k sqft (Size Group 6)***

Assuming median occupancy, median rent = $32/sqft

Diff between non-green and green class A in this size group = $32-$29 =$3/sqft

=> Minimum additional revenue per year = 3*250000 = $750,000 (Ensures no addditional cost incurred in baseline construction)

Given that we spend an extra 5 million dollars for the green building,with a median occupancy rate of 93.9% for 250,000 sqft green, class A building, we will be able to recuperate this cost by 7.099 years. 


But assuming an occupancy rate of ~94% might be far from reality, as the residential real estate industry thrives from the word-of-mouth and reviews of existing/old tenants. Hence, we take a level based approach of assuming 80% occupancy for first 3 years and 90% from year 3 onward and also, we consider a conservative estimate of 15% inflation rate for difference in rent after 3 years.  

This implies, in first 3 years...

Additional Revenue = 3*250000*0.8 = $600,000/year

After the first 3 years, we expect the diff in rent/per sqft to go up by 15% and occupancy to be 90%.

Additional Revenue= 3.45*250000*0.9= $776,250/year


Even with these conservative assumptions for occupancy rate, we will be able recuperate the costs in 7.123 years.  

