---
title: "Association Analysis"
output:
  html_document:
    df_print: paged
date: "2023-08-14"
---
Anisha Singh (as229583), Antra Sinha (as224589), Hayoung Kim (hk26786), Meenakshi Sundaram (ms88544)

Association rule mining
Revisit the notes on association rule mining and the R example on music playlists: playlists.R and playlists.csv. Then use the data on grocery purchases in groceries.txt and find some interesting association rules for these shopping baskets. The data file is a list of shopping baskets: one person's basket for each row, with multiple items per row separated by commas. Pick your own thresholds for lift and confidence; just be clear what these thresholds are and say why you picked them. Do your discovered item sets make sense? Present your discoveries in an interesting and visually appealing way.

Notes:

This is an exercise in visual and numerical story-telling. Do be clear in your description of what you've done, but keep the focus on the data, the figures, and the insights your analysis has drawn from the data, rather than technical details.
The data file is a list of baskets: one row per basket, with multiple items per row separated by commas. You'll have to cobble together your own code for processing this into the format expected by the "arules" package. This is not intrinsically all that hard, but it is the kind of data-wrangling wrinkle you'll encounter frequently on real problems, where your software package expects data in one format and the data comes in a different format. Figuring out how to bridge that gap is part of the assignment, and so we won't be giving tips on this front.


```{r setup, include=FALSE}
#install.packages("arules")
library(arules)
library(arulesViz)
```

```{r}
# Load the data from groceries.txt
data <- readLines("groceries.txt")
# Process the data into transactions format
transactions <- strsplit(data, split = ",") 
```

```{r}
# Convert the transactions into transactions class (required for arules)
transactions <- as(transactions, "transactions")
summary(transactions)
```
# 1. Top 20 frequently purchased items

```{r}
# Create a scatter plot of item frequency
itemFrequencyPlot(transactions, topN = 20, main= 'top20 frequently purchased items')
```

- Whole milk, other vegetables, rolls/buns, soda, and yogurt are the top 5 most frequently purchased items.
- Daily ingredients or dairy products are most frequently bought.


# 2. Range of support, confidence, and lift
```{r}
# Perform association rule mining using the apriori algorithm without filtering.
rules <- apriori(transactions, parameter = list(supp = 0.01, conf = 0.2, target = "rules", minlen = 2))
summary(rules)

```

Apply the Apriori Algorithm for the first, without filtering 
From the summary of rules drawn out from Apriori algorithm, we can know the ranges of each parameter

- Possible Support range: from 0.01007 to 0.07483
- Possible Confidence range: from 0.2006 to 0.5862
- Possible Lift range: from 0.8991 to 3.2950

# 3. Plot of support,lift,confidence
```{r}
plot(rules, measure = c("support", "lift"), shading = "confidence")
```


# 3-1. "two key" plot
```{r}
# "two key" plot: colors represent the lift
plot(rules, method='two-key plot', main = 'Scatterplot of 231 rules, with color variations by lift 2 or lift 3')

```



# 4. Plot of entire rules before filtering (total 231 rules)
```{r}
#Plotting associations before filtering
plot(rules, method = "graph", control = list(type = "items"))
```

- A bunch of meaningless rules that are hard to recognize, so we chose to find meaningful subsets by filtering.Let's apply filter and gain some insights.


# 5. Applying filter with lift>2 and confidence>0.3 (total 44 rules)
```{r}
# Filter the rules based on lift and confidence thresholds
interesting_rules <- subset(rules, lift > 2 & confidence > 0.3)
# Show the discovered association rules
inspect(interesting_rules)

```

# 5-1. Finding and Plotting the associations of difference rules based on confidence,lift,support (total 44 rules)
```{r}
# Extract the relevant measures from the rules
support <- interestMeasure(interesting_rules, "support")
confidence <- interestMeasure(interesting_rules, "confidence")
lift <- interestMeasure(interesting_rules, "lift")

# Create a data frame for the measures
measures_df <- data.frame(Support = support, Confidence = confidence, Lift = lift)

```

```{r fig2, echo=FALSE, fig.align='left', fig.height = 10, fig.width = 7}
# Create a bar chart of different association rule measures
barplot(t(as.matrix(measures_df)), beside = TRUE, col = c("blue", "green", "orange"), main = "Association Rule Measures",
        xlab = "Rules", ylab = "Measures", legend.text = rownames(measures_df), args.legend = list(x = "topright"))
```

- By examining the heights of the bars, we tried to quickly identify which rules have higher support, confidence, and lift and therefore might be more significant or interesting for the given dataset. 

- The rules represented in orange clearly appear to be significant, and the idea of narrowing down to approximately 10 rules for further consideration has been gained.


# 5-2. Plot of rules after filtering (total 44 rules)
```{r}
#Plotting association after filtering
plot(interesting_rules, method = "graph", control = list(type = "items"))
```


# 6. Inspect the top 10 rules by confidence
```{r}
# Inspect the top 10 rules by confidence
top_rules_by_confidence <- head(sort(interesting_rules, by = "confidence"), n = 10)
inspect(top_rules_by_confidence)
```

### Findings
### - Strong Association with Other Vegetables <analysis 1>:
We can see some strong confidence value such as {citrus fruit, root vegetables},{root vegetables, tropical fruit},{curd, yogurt},{butter, other vegetables},{root vegetables, tropical fruit} with value over 0.56.Therefore the first two rules highlight a strong association between combinations of fruits and vegetables (citrus fruit & root vegetables, root vegetables & tropical fruit) and the purchase of other vegetables.


### - Prominent Role of Whole Milk <analysis 2> : 
Rules 3 to 10 all involve the purchase of whole milk. This suggests that whole milk is a common item that is frequently bought in conjunction with various other products, such as curd & yogurt, butter & other vegetables, root vegetables & tropical fruit, and more.



# 7. Inspect the top 10 rules by lift
```{r}
# Inspect the top rules by lift
top_rules_by_lift <- head(sort(interesting_rules, by = "lift"), n = 10)
inspect(top_rules_by_lift)
```
### Findings
- Root Vegetables Connection: 
A significant pattern is the frequent association of various products with root vegetables. This occurs with combinations involving citrus fruit, other vegetables, tropical fruit, and even beef. Marketing strategies could capitalize on this by bundling root vegetables with these products or positioning them close together in the store.

- Other Vegetables Association: 
Several rules indicate that other vegetables are commonly bought with various combinations of products, including citrus fruit, root vegetables, and tropical fruit. This could imply a general trend toward purchasing vegetables, possibly reflecting health-conscious shopping habits.


## The best rule - which has the highest lift of 3.295  <analysis 3>

- The rule {citrus fruit, other vegetables} => {root vegetables} reveals that when customers buy citrus fruit and other vegetables, they also tend to buy root vegetables.

- The lift value of 3.295 indicates that the likelihood of purchasing root vegetables is 3.295 times higher when citrus fruit and other vegetables are present in the basket. 


# 8. Subset of "lift> 2.5 & confidence > 0.4" which filtered into 5 rules.

```{r}
# Filter the rules based on lift and confidence thresholds
interesting_rules <- subset(rules, lift > 2.5 & confidence > 0.4)
# Show the discovered association rules
inspect(interesting_rules)

#Plotting after filtering association
plot(interesting_rules, method = "graph", control = list(type = "items"))

```

### Findings <analysis 4>
- Strong Relationship with Other Vegetables: 
Each rule in this set highlights a strong association between certain combinations of items and the purchase of other vegetables. The rules demonstrate various combinations of items, such as whipped/sour cream & yogurt, citrus fruit & root vegetables, root vegetables & tropical fruit, etc., leading to the purchase of other vegetables.

- Root Vegetables Commonly Paired: 
Root vegetables appear in four of the five rules, suggesting that they are commonly paired with other items such as citrus fruit, tropical fruit, yogurt, and rolls/buns. This could indicate a particular preference or trend among customers for combining root vegetables with various other products.



# 9. Conclusion:
- Strong Association with Other Vegetables: 
In <analysis 1 and 2>, we see a consistent pattern where various combinations of products are associated with the purchase of other vegetables. This includes combinations like citrus fruit & root vegetables, root vegetables & tropical fruit, whipped/sour cream & yogurt, etc.

- Whole Milk as a Common Item: 
In <analysis 2>, there is a noticeable trend where different combinations of products are associated with the purchase of whole milk.

- Specific Combinations with High Lift: 
The high lift values in many rules, such as {citrus fruit, root vegetables} => {other vegetables}, indicate a strong association beyond mere coincidence.

- Different Products Leading to Similar Purchases: 
<Analysis 3 and 4> show some overlap with the combination of citrus fruit and root vegetables leading to other vegetables in one case and root vegetables in another.


