---
title: "Association Analysis"
output:
  html_document:
    df_print: paged
date: "2023-08-14"
---

```{r setup, include=FALSE}
#install.packages("arules")
library(arules)
library(arulesViz)
```

```{r}
# Load the data from groceries.txt
data <- readLines("groceries.txt")
# Process the data into transactions format
transactions <- strsplit(data, split = ",") 
```

```{r}
# Convert the transactions into transactions class (required for arules)
transactions <- as(transactions, "transactions")
summary(transactions)
```
# 1. Top 20 frequently purchased items
```{r}
# Create a scatter plot of item frequency
itemFrequencyPlot(transactions, topN = 20, main= 'top20 frequently purchased items')
```
# 2. Range of support, confidence, and lift
```{r}
# Perform association rule mining using the apriori algorithm without filtering.
rules <- apriori(transactions, parameter = list(supp = 0.01, conf = 0.2, target = "rules", minlen = 2))
summary(rules)

```

Apply the Apriori Algorithm for the first, without filtering 
From the summary of rules drawn out from Apriori algorithm, we can know the ranges of each parameter

- Possible Support range: from 0.01007 to 0.07483
- Possible Confidence range: from 0.2006 to 0.5862
- Possible Lift range: from 0.8991 to 3.2950

# 3. Plot of support,lift,confidence
```{r}
plot(rules, measure = c("support", "lift"), shading = "confidence")
```
# 4."two key" plot
```{r}
# "two key" plot: coloring is by size (order) of item set
plot(rules, method='two-key plot')

```



# 5.Plot of entire rules before filtering 
```{r}
#Plotting associations before filtering
plot(rules, method = "graph", control = list(type = "items"))
```
- A bunch of meaningless rules that are hard to recognize, so we chose to find meaningful subsets by filtering.Let's apply filter and gain some insights.


# 5.Applying filter with lift>2 and confidence>0.3
```{r}
# Filter the rules based on lift and confidence thresholds
interesting_rules <- subset(rules, lift > 2 & confidence > 0.3)
# Show the discovered association rules
inspect(interesting_rules)

```

# 6.Finding and Plotting the associations of difference rules based on confidence,lift,support
```{r}
# Extract the relevant measures from the rules
support <- interestMeasure(interesting_rules, "support")
confidence <- interestMeasure(interesting_rules, "confidence")
lift <- interestMeasure(interesting_rules, "lift")

# Create a data frame for the measures
measures_df <- data.frame(Support = support, Confidence = confidence, Lift = lift)

```

```{r}
# Create a bar chart of different association rule measures
barplot(t(as.matrix(measures_df)), beside = TRUE, col = c("blue", "green", "orange"), main = "Association Rule Measures",
        xlab = "Rules", ylab = "Measures", legend.text = rownames(measures_df), args.legend = list(x = "topright"))
```

# 7. Plot of entire rules after filtering 
```{r}
#Plotting association after filtering
plot(interesting_rules, method = "graph", control = list(type = "items"))
```

# 8. Inspect the top 10 rules by confidence
```{r}
# Inspect the top 10 rules by confidence
top_rules_by_confidence <- head(sort(interesting_rules, by = "confidence"), n = 10)
inspect(top_rules_by_confidence)
```
#Findings

- Strong Association with Other Vegetables: <analysis 1>
#We can see some strong confidence value such as {citrus fruit, root vegetables},{root vegetables, tropical fruit},{curd, yogurt},{butter, other vegetables},{root vegetables, tropical fruit} with value over 0.56.Therefore the first two rules highlight a strong association between combinations of fruits and vegetables (citrus fruit & root vegetables, root vegetables & tropical fruit) and the purchase of other vegetables.


- Prominent Role of Whole Milk: <analysis 2>
#Rules 3 to 10 all involve the purchase of whole milk. This suggests that whole milk is a common item that is frequently bought in conjunction with various other products, such as curd & yogurt, butter & other vegetables, root vegetables & tropical fruit, and more.



# 9. Inspect the top 10 rules by lift
```{r}
# Inspect the top rules by lift
top_rules_by_lift <- head(sort(interesting_rules, by = "lift"), n = 10)
inspect(top_rules_by_lift)
```

#The best rule - which has the highest lift of 3.295 <analysis 3>
### Finding
- The rule {citrus fruit, other vegetables} => {root vegetables} reveals that when customers buy citrus fruit and other vegetables, they also tend to buy root vegetables.

- The lift value of 3.295 indicates that the likelihood of purchasing root vegetables is 3.295 times higher when citrus fruit and other vegetables are present in the basket. 


#10.Subset of "lift> 2.5 & confidence > 0.4" which filtered into 5 rules.

```{r}
# Filter the rules based on lift and confidence thresholds
interesting_rules <- subset(rules, lift > 2.5 & confidence > 0.4)
# Show the discovered association rules
inspect(interesting_rules)

#Plotting after filtering association
plot(interesting_rules, method = "graph", control = list(type = "items"))

```


# Subset of "lift> 2.5 & confidence > 0.4" which filtered into 5 rules <analysis 4>
### Findings
- Strong Relationship with Other Vegetables: 
Each rule in this set highlights a strong association between certain combinations of items and the purchase of other vegetables. The rules demonstrate various combinations of items, such as whipped/sour cream & yogurt, citrus fruit & root vegetables, root vegetables & tropical fruit, etc., leading to the purchase of other vegetables.

- Root Vegetables Commonly Paired: 
Root vegetables appear in four of the five rules, suggesting that they are commonly paired with other items such as citrus fruit, tropical fruit, yogurt, and rolls/buns. This could indicate a particular preference or trend among customers for combining root vegetables with various other products.



## Conclusion:
- Strong Association with Other Vegetables: 
In <analysis 1 and 2>, we see a consistent pattern where various combinations of products are associated with the purchase of other vegetables. This includes combinations like citrus fruit & root vegetables, root vegetables & tropical fruit, whipped/sour cream & yogurt, etc.

- Whole Milk as a Common Item: 
In <analysis 1>, there is a noticeable trend where different combinations of products are associated with the purchase of whole milk.

- Specific Combinations with High Lift: 
The high lift values in many rules, such as {citrus fruit, root vegetables} => {other vegetables}, indicate a strong association beyond mere coincidence.

- Different Products Leading to Similar Purchases: 
<Analysis 3 and 4> show some overlap with the combination of citrus fruit and root vegetables leading to other vegetables in one case and root vegetables in another.



